\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2015-05-21 doc/USER_MANUAL/03_mesh_generation.tex
updates manual for better wiki display

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.6\textwidth]{figures/mount-cubit.jpg}
\par\end{centering}

\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
shown in the main display consists of a hexahedral discretization
of a single volume with topography.}

\label{fig:mount.cubit}
\end{figure}

@@

For the installation and handling of the CUBIT meshing tool suite,
please refer to the CUBIT user manual and documentation. In order
to give you a basic understanding of how to use CUBIT for our purposes,
examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
volumes and creates a refinement layer between the two. This example
can be compared for validation against the solutions provided in subdirectory
~\\
 \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose
material properties will have to be read in from a tomographic model
file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/example-homogeneous.jpg}
\includegraphics[width=0.45\textwidth]{figures/example-2layers.jpg} \\
\includegraphics[width=0.45\textwidth]{figures/example-water.jpg}
\includegraphics[width=0.45\textwidth]{figures/example-tomo.jpg}
\par\end{centering}

\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
homogeneous halfspace (top-left), layered halfspace (top-right), water
layered halfspace (bottom-left) and tomographic model (bottom-right).}

\label{fig:examples.cubit}
\end{figure}

@@

For an acoustic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{0}] S-wave speed of the material is ignored.
\item [{rho}] density of the material (given in kg/m$^{3}$).
\end{description}
\item [{face\_topo}] Block definition for the surface which defines the
free surface (which can have topography). The name of this block must
be 'face\_topo', the block has to be defined using all the surfaces
which constitute the complete free surface of the model.
\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmin.
\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmax.
\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymin.
\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymax.
\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
boundaries, one block for each surface with z=bottom.
\end{description}
\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-surface.jpg}
\includegraphics[width=0.45\textwidth]{figures/mount-abs.jpg}
\par\end{centering}

\caption{Example of the block definitions for the free surface 'face\_topo'
(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
(right) in CUBIT.}

\label{fig:mount.abs}
\end{figure}

@@

In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:
{\small
\begin{verbatim}
    0 1 vol1 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
    0 2 vol2 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
\end{verbatim}
}
(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance
{\small
\begin{verbatim}
    2 1 1500 2300 1800 9999.0 9999.0 0
    2 2 1600 2500 20000 9999.0 9999.0 0
\end{verbatim}
}

@@

An efficient partitioning is important because it leverages the overall
running time of the application. It amounts to balance the number
of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors.
\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning
routines. SCOTCH is a free software package developed by Fran\c{c}ois
Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
from the web page \url{https://gforge.inria.fr/projects/scotch/}.
\\

In most cases, the configuration with \texttt{./configure~FC=ifort}
should be sufficient. During the configuration process, the script
tries to find existing SCOTCH installations. In case your system has
no pre-existing SCOTCH installation, we provide the source code of
SCOTCH, which is released open source under the French CeCILL-C version
1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
This version gets bundled with the compilation of the SPECFEM3D Cartesian
package if no libraries could have been found. If this automatic compilation
of the SCOTCH libraries fails, please refer to file INSTALL.txt in
that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you
have correctly specified the path of the SCOTCH library when using
the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
script. In the future you should be able to find more recent versions
at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}
. \\

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-partitions.jpg}
\includegraphics[width=0.45\textwidth]{figures/mount-partitions2.jpg}
\par\end{centering}

\caption{Example of a mesh partitioning onto four cores. Each single core partition
is colored differently. The executable \texttt{xdecompose\_mesh} can
equally distribute the mesh on any arbitrary number of cores. Domain
decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
scaling up to 150,000 processor cores in shown for instance in \citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}

\label{fig:mount.partitions}
\end{figure}

When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
make xdecompose_mesh
\end{verbatim}
}
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. The synopsis is:
\begin{verbatim}
./bin/xdecompose_mesh nparts input_directory output_directory
\end{verbatim}
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

@@

In case you successfully ran the configuration script, you are also
ready to compile the internal mesher. This is an alternative to CUBIT
for the mesh generation of relatively simple models.

\begin{figure}[htbp]
\noindent \begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi.pdf}
\par\end{centering}

\caption{For parallel computing purposes,
the model block is subdivided in $\nprocxi\times\nproceta$ slices
of elements. In this example we use $5^{2}=25$ processors. }
\label{fig:For-parallel-computing}
\end{figure}

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xmeshfem3D
\end{verbatim}
}
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

@@

The mesher and the solver use UTM coordinates internally, therefore
you need to define the zone number for the UTM projection (e.g., zone
11 for Los Angeles). Use decimal values for latitude and longitude
(no minutes/seconds). These values are approximate; the mesher will
round them off to define a square mesh in UTM coordinates. When running
benchmarks on rectangular models, turn the UTM projection off by using
the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}{\small , in which
case all `longitude' parameters simply refer to the $x$~axis, and
all `latitude' parameters simply refer to the $y$~axis. To run the
mesher for a global simulation, the following parameters need to be
set in the }\texttt{\small Mesh\_Par\_file}{\small :}{\small \par}
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{\noun{UTM\_PROJECTION\_ZONE}}}] UTM projection zone in
which your model resides, only valid when \texttt{SUPPRESS\_UTM\_}~\\
 \texttt{PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.
We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.
From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\\
The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
which means that current UTM northing at a given point can be 200+ meters different from the old one.
For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.
%
\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
\item [{$\nexxi$}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
\end{equation}
 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
\item [{$\nexeta$}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
\item [{$\nprocxi$}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
\item [{$\nproceta$}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
is set to \texttt{.true.}.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}).
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).
The mesher generates the necessary partitions in parallel, one set
for each of the $\nprocxi\times\nproceta$ slices that constitutes
the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
mesher finishes, you can log in to one of the compute nodes and view
the contents of the \texttt{LOCAL\_PATH} directory to see the files
generated by the mesher. These files will be needed for creating the
distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\item [{\texttt{NMATERIALS}}] The number of different materials in your
model. In the following lines, each material needs to be defined as
:

@@

where
\begin{itemize}
\item \texttt{Q} : quality factor (0=no attenuation)
\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
in \texttt{aniso\_model.f90}
\item \texttt{domain\_id} : 1=acoustic / 2=elastic
\end{itemize}
\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
lines, because the mesh is regular or 'almost regular', each region
is defined as :

>>>>

\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.6\textwidth]{figures/mount-cubit.jpg}
\par
\end{centering}
\caption{Example of the graphical user interface of CUBIT. The hexahedral mesh
shown in the main display consists of a hexahedral discretization
of a single volume with topography.}
\label{fig:mount.cubit}
\end{figure}

@@

For the installation and handling of the CUBIT meshing tool suite,
please refer to the CUBIT user manual and documentation. In order
to give you a basic understanding of how to use CUBIT for our purposes,
examples are provided in the SPECFEM3D Cartesian package in the subdirectory
\texttt{EXAMPLES/}:
\begin{description}
\item [{\texttt{homogeneous\_halfspace}}] Creates a single block model
and assigns elastic material parameters.
\item [{\texttt{layered\_halfspace}}] Combines two different, elastic material
volumes and creates a refinement layer between the two. This example
can be compared for validation against the solutions provided in subdirectory
~\\
 \texttt{VALIDATION\_3D\_SEM\_SIMPLER\_LAYER\_SOURCE\_DEPTH/}.
\item [{\texttt{waterlayered\_halfspace}}] Combines an acoustic and elastic
material volume as in a schematic marine survey example.
\item [{\texttt{tomographic\_model}}] Creates a single block model whose
material properties will have to be read in from a tomographic model
file during the databases creation by \texttt{xgenerate\_databases}.
\end{description}
%
\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.45\textwidth]{figures/example-homogeneous.jpg}
\includegraphics[width=0.45\textwidth]{figures/example-2layers.jpg} \\
\includegraphics[width=0.45\textwidth]{figures/example-water.jpg}
\includegraphics[width=0.45\textwidth]{figures/example-tomo.jpg}
\par
\end{centering}
\caption{Screenshots of the CUBIT examples provided in subdirectory \texttt{EXAMPLES/}:
homogeneous halfspace (top-left), layered halfspace (top-right), water
layered halfspace (bottom-left) and tomographic model (bottom-right).}
\label{fig:examples.cubit}
\end{figure}

@@

For an acoustic material:
\begin{description}
\item [{material\_id}] An integer value which is unique for this material.
\item [{Vp}] P-wave speed of the material (given in m/s).
\item [{0}] S-wave speed of the material is ignored.
\item [{rho}] density of the material (given in kg/m$^{3}$).
\end{description}
\item [{face\_topo}] Block definition for the surface which defines the
free surface (which can have topography). The name of this block must
be 'face\_topo', the block has to be defined using all the surfaces
which constitute the complete free surface of the model.
\item [{face\_abs\_xmin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmin.
\item [{face\_abs\_xmax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with x=Xmax.
\item [{face\_abs\_ymin}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymin.
\item [{face\_abs\_ymax}] Block definition for the faces on the absorbing
boundaries, one block for each surface with y=Ymax.
\item [{face\_abs\_bottom}] Block definition for the faces on the absorbing
boundaries, one block for each surface with z=bottom.
\end{description}
%
\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-surface.jpg}
\includegraphics[width=0.45\textwidth]{figures/mount-abs.jpg}
\par
\end{centering}
\caption{Example of the block definitions for the free surface 'face\_topo'
(left) and the absorbing boundaries, defined in a single block 'face\_abs\_xmin'
(right) in CUBIT.}
\label{fig:mount.abs}
\end{figure}

@@

In directory "CUBIT\_GEOCUBIT" we provide a script that can help doing the above tasks of exporting a CUBIT mesh
to SPECFEM3D format automatically for you: "run\_cubit2specfem3d.py".
Just edit them to indicate the path to your local installation of CUBIT and also the name of the *.cub existing CUBIT mesh file
that you want to export to SPECFEM3D format. These scripts will do the conversion for you automatically except assigning material properties
to the different mesh layers. To do so, you will then need to edit the file
called "nummaterial\_velocity\_file" that will have just been created and change it from the prototype created:

{\footnotesize
\begin{verbatim}
0 1 vol1 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
0 2 vol2 --> syntax: #material_domain_id #material_id #rho #vp #vs #Q_kappa #Q_mu #anisotropy
\end{verbatim}
}

(where "vol1" and "vol2" here represent the volume labels that you have set while creating the mesh in CUBIT)
to for instance

{\footnotesize
\begin{verbatim}
2 1 1500 2300 1800 9999.0 9999.0 0
2 2 1600 2500 20000 9999.0 9999.0 0
\end{verbatim}
}

@@

An efficient partitioning is important because it leverages the overall
running time of the application. It amounts to balance the number
of elements in each slice while minimizing the communication costs
resulting from the placement of adjacent elements on different processors.
\texttt{decompose\_mesh} depends on the SCOTCH library \citep{PeRo96},
which provides efficent static mapping, graph and mesh partitioning
routines. SCOTCH is a free software package developed by Fran\c{c}ois
Pellegrini et al. from LaBRI and INRIA in Bordeaux, France, downloadable
from the web page \url{https://gforge.inria.fr/projects/scotch/}.\\

In most cases, the configuration with \texttt{./configure~FC=ifort}
should be sufficient. During the configuration process, the script
tries to find existing SCOTCH installations. In case your system has
no pre-existing SCOTCH installation, we provide the source code of
SCOTCH, which is released open source under the French CeCILL-C version
1 license, in directory \texttt{src/decompose\_mesh/scotch\_5.1.12b}.
This version gets bundled with the compilation of the SPECFEM3D Cartesian
package if no libraries could have been found. If this automatic compilation
of the SCOTCH libraries fails, please refer to file INSTALL.txt in
that directory to see further details how to compile it on your system.
In case you want to use a pre-existing installation, make sure you
have correctly specified the path of the SCOTCH library when using
the option \texttt{-{}-with-scotch-dir} with the \texttt{./configure}
script. In the future you should be able to find more recent versions
at \url{http://www.labri.fr/perso/pelegrin/scotch/scotch_en.html}.\\

\begin{figure}[htbp]
\begin{centering}
\includegraphics[width=0.45\textwidth]{figures/mount-partitions.jpg}
\includegraphics[width=0.45\textwidth]{figures/mount-partitions2.jpg}
\par
\end{centering}
\caption{Example of a mesh partitioning onto four cores. Each single core partition
is colored differently. The executable \texttt{xdecompose\_mesh} can
equally distribute the mesh on any arbitrary number of cores. Domain
decomposition is explained in detail in \citet{MaKoBlLe08}, and excellent
scaling up to 150,000 processor cores in shown for instance in 
\citet{CaKoLaTiMiLeSnTr08,KoLaMi08a,MaKoBlLe08,KoErGoMi10,Kom11}.}
\label{fig:mount.partitions}
\end{figure}

When you are ready to compile, in the main directory type
{\small
\begin{verbatim}
make xdecompose_mesh
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the executable \texttt{bin/xdecompose\_mesh}
should be produced.

The partitioning is done in serial for now (in the next release we
will provide a parallel version of that code). It needs to be run in the main directory because it expects the \texttt{./DATA/Par\_file}. 
The synopsis is:
%
\begin{verbatim}
./bin/xdecompose_mesh nparts input_directory output_directory
\end{verbatim}
%
\noindent
where
\begin{itemize}
\item \texttt{nparts} is the number of partitions, i.e., the number of cores
for the parallel simulations,
\item \texttt{input\_directory} is the directory which holds all the files
generated by the Python script \texttt{cubit2specfem3d.py} explained
in the previous Section~\ref{subsec:Exporting-the-Mesh}, e.g. \texttt{./MESH/},
and
\item \texttt{output\_directory} is the directory for the output of this
partitioner which stores ACII-format files named like \texttt{proc{*}{*}{*}{*}{*}{*}\_Database}
for each partition. These files will be needed for creating the distributed
databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{./OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
\end{itemize}
Note that all the files generated by the Python script \texttt{cubit2specfem3d.py}
must be placed in the \texttt{input\_directory} folder before running
the program.

@@

In case you successfully ran the configuration script, you are also
ready to compile the internal mesher. This is an alternative to CUBIT
for the mesh generation of relatively simple models.
%
\begin{figure}[htbp]
\begin{centering}
\includegraphics[scale=0.5]{figures/socal_map_mpi.pdf}
\par
\end{centering}
\caption{For parallel computing purposes,
the model block is subdivided in $\nprocxi\times\nproceta$ slices
of elements. In this example we use $5^{2}=25$ processors. }
\label{fig:For-parallel-computing}
\end{figure}

\noindent
In the main directory, type
{\small
\begin{verbatim}
make xmeshfem3D
\end{verbatim}
}
\noindent
If all paths and flags have been set correctly, the mesher should
now compile and produce the executable \texttt{bin/xmeshfem3D}. Please
note that \texttt{xmeshfem3D} must be called directly from the main
directory, as most of the binaries of the package.

@@

The mesher and the solver use UTM coordinates internally, therefore
you need to define the zone number for the UTM projection (e.g., zone
11 for Los Angeles). Use decimal values for latitude and longitude
(no minutes/seconds). These values are approximate; the mesher will
round them off to define a square mesh in UTM coordinates. When running
benchmarks on rectangular models, turn the UTM projection off by using
the flag \texttt{\small SUPPRESS\_UTM\_PROJECTION}, 
in which case all `longitude' parameters simply refer to the $x$~axis, and
all `latitude' parameters simply refer to the $y$~axis. To run the
mesher for a global simulation, the following parameters need to be
set in the \texttt{\small Mesh\_Par\_file}:
%
\begin{description}
\item [{\texttt{LATITUDE\_MIN}}] Minimum latitude in the block (negative
for South).
\item [{\texttt{LATITUDE\_MAX}}] Maximum latitude in the block.
\item [{\texttt{LONGITUDE\_MIN}}] Minimum longitude in the block (negative
for West).
\item [{\texttt{LONGITUDE\_MAX}}] Maximum longitude in the block.
\item [{\texttt{DEPTH\_BLOCK\_KM}}] Depth of bottom of mesh in kilometers.
\item [{\texttt{UTM\_PROJECTION\_ZONE}}] UTM projection zone in
which your model resides, only valid when 
\texttt{SUPPRESS\_UTM\_PROJECTION} is \texttt{.false.}.
Use a negative zone number for the Southern hemisphere:
the Northern hemisphere corresponds to zones +1 to +60,
the Southern hemisphere to zones -1 to -60.

We use the WGS84 (World Geodetic System 1984) reference ellipsoid for the UTM projection. If you prefer to use the Clarke 1866 ellipsoid,
edit file \texttt{src/shared/utm\_geo.f90}, uncomment that ellipsoid and recompile the code.

From \url{http://en.wikipedia.org/wiki/Universal_Transverse_Mercator_coordinate_system}:\\
The Universal Transverse Mercator coordinate system was developed by the United States Army Corps of Engineers in the 1940s.
The system was based on an ellipsoidal model of Earth. For areas within the contiguous United States
the Clarke Ellipsoid of 1866 was used. For the remaining areas of Earth, including Hawaii, the International Ellipsoid was used.
The WGS84 ellipsoid is now generally used to model the Earth in the UTM coordinate system,
which means that current UTM northing at a given point can be 200+ meters different from the old one.
For different geographic regions, other datum systems (e.g.: ED50, NAD83) can be used.

\item [{\texttt{SUPPRESS\_UTM\_PROJECTION}}] set to be \texttt{.false.}
when your model range is specified in geographical coordinates, and
needs to be \texttt{.true.} when your model is specified in Cartesian
coordinates. \noun{UTM projection zone in which your simulation region
resides.}
%
\item [{\texttt{INTERFACES\_FILE }}] File which contains the description
of the topography and of the interfaces between the different layers
of the model, if any. The number of spectral elements in the vertical
direction within each layer is also defined in this file.
%
\item [{\texttt{NEX\_XI}}] The number of spectral elements along one side of the
block. This number \textit{must} be 8~$\times$~a multiple of $\nprocxi$
defined below. Based upon benchmarks against semi-analytical discrete
wavenumber synthetic seismograms \citep{KoLiTrSuStSh04}, determined
that a $\nexxi=288$ run is accurate to a shortest period of roughly
2~s. Therefore, since accuracy is determined by the number of grid
points per shortest wavelength, for any particular value of $\nexxi$
the simulation will be accurate to a shortest period determined by
\begin{equation}
\mbox{shortest period (s)}=(288/\nexxi)\times2.\label{eq:shortest_period}
\end{equation}
 The number of grid points in each orthogonal direction of the reference
element, i.e., the number of Gauss-Lobatto-Legendre points, is determined
by \texttt{NGLLX} in the \texttt{constants.h} file. We generally use
$\mbox{\texttt{NGLLX\/}}=5$, for a total of $5^{3}=125$ points per
elements. We suggest not to change this value.
%
\item [{\texttt{NEX\_ETA}}] The number of spectral elements along the other side
of the block. This number \textit{must} be 8~$\times$~a multiple
of $\nproceta$ defined below.
%
\item [{\texttt{NPROC\_XI}}] The number of processors or slices along one side
of the block (see Figure~\ref{fig:For-parallel-computing}); we must
have $\nexxi=8\times c\times\nprocxi$, where $c\ge1$ is a positive
integer.
%
\item [{\texttt{NPROC\_ETA}}] The number of processors or slices along the other
side of the block; we must have $\nexeta=8\times c\times\nproceta$,
where $c\ge1$ is a positive integer.
%
\item [{\texttt{USE\_REGULAR\_MESH}}] set to be \texttt{.true.} if you
want a perfectly regular mesh or \texttt{.false.} if you want to add
doubling horizontal layers to coarsen the mesh. In this case, you
also need to provide additional information by setting up the next
three parameters.
%
\item [{\texttt{NDOUBLINGS}}] The number of horizontal doubling layers.
Must be set to \texttt{1} or \texttt{2} if \texttt{USE\_REGULAR\_MESH}
is set to \texttt{.true.}.
\item [{\texttt{NZ\_DOUBLING\_1}}] The position of the first doubling layer
(only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}).
\item [{\texttt{NZ\_DOUBLING\_2}}] The position of the second doubling
layer (only interpreted if \texttt{USE\_REGULAR\_MESH} is set to \texttt{.true.}
and if \texttt{NDOUBLINGS} is set to \texttt{2}).
\item [{\texttt{CREATE\_ABAQUS\_FILES}}] Set this flag to \texttt{.true.}
to save Abaqus FEA \urlwithparentheses{www.simulia.com} mesh files
for subsequent viewing. Turning the flag on generates files in the
\texttt{LOCAL\_PATH} directory. See Section~\ref{sec:Mesh-graphics}
for a discussion of mesh viewing features.
\item [{\texttt{CREATE\_DX\_FILES}}] Set this flag to \texttt{.true.} to
save OpenDX \urlwithparentheses{www.opendx.org} mesh files for subsequent
viewing.
\item [{\texttt{LOCAL\_PATH}}] Directory in which the partitions generated
by the mesher will be written. Generally one uses a directory on the
local disk of the compute nodes, although on some machines these partitions
are written on a parallel (global) file system (see also the earlier
discussion of the \texttt{LOCAL\_PATH\_IS\_ALSO\_GLOBAL} flag in Chapter~\ref{cha:Getting-Started}).

The mesher generates the necessary partitions in parallel, one set
for each of the $\nprocxi\times\nproceta$ slices that constitutes
the mesh (see Figure~\ref{fig:For-parallel-computing}). After the
mesher finishes, you can log in to one of the compute nodes and view
the contents of the \texttt{LOCAL\_PATH} directory to see the files
generated by the mesher. These files will be needed for creating the
distributed databases, and have to reside in the directory \texttt{LOCAL\_PATH}
specified in the main \texttt{Par\_file}, e.g. in directory \texttt{OUTPUT\_FILES/DATABASES\_MPI}.
Please see Chapter~\ref{cha:Creating-Distributed-Databases} for
further details.
%
\item [{\texttt{NMATERIALS}}] The number of different materials in your
model. In the following lines, each material needs to be defined as:

@@

\noindent
where
\begin{itemize}
\item \texttt{Q} : quality factor (0=no attenuation)
\item \texttt{anisotropy\_flag} : 0=no anisotropy / 1,2,.. check with implementation
in \texttt{aniso\_model.f90}
\item \texttt{domain\_id} : 1=acoustic / 2=elastic
\end{itemize}
\item [{\texttt{NREGIONS}}] The number of regions in the mesh. In the following
lines, because the mesh is regular or 'almost regular', each region
is defined as:
\begin{verbatim}
NEX_XI_BEGIN NEX_XI_END NEX_ETA_BEGIN NEX_ETA_END NZ_BEGIN NZ_END material_ID
\end{verbatim}
\end{description}
%
The \texttt{INTERFACES\_FILE} parameter of \texttt{Mesh\_Par\_File}
defines the file which contains the settings of the topography grid
and of the interfaces grids. Topography is defined as a set of elevation
values on a regular 2D grid. It is also possible to define interfaces
between the layers of the model in the same way. The file needs to
define several parameters:
\begin{itemize}
\item The number of interfaces, including the topography. This needs to
be set at the first line. Then, from the bottom to the top of the
model, you need to define the grids with:
\item \texttt{SUPPRESS\_UTM\_PROJECTION} flag as described previously,
\item number of points along $x$ and $y$ direction (NXI and NETA),
\item minimal $x$ and $y$ coordinates (LONG\_MIN and LAT\_MIN),
\item spacing between points along $x$ and $y$ (SPACING\_XI and SPACING\_ETA)
and
\item the name of the file which contains the elevation values (in $y$.$x$
increasing order).
\end{itemize}
%
At the end of this file, you simply need to set the number of spectral
elements in the vertical direction for each layer. We provide a few
models in the {\texttt{EXAMPLES/}} directory.\\
